### 2023-06-05 proofs/tex/unification-lemmas.tex
WIP: finish merge lemmas

        By the soundness of the unification solution merge (\cref{lemma:unif-sol-merge-soundness}), we have:
        \begin{itemize}
            \item $[[us1 & us2 : Θ | uv uP ∪ uv uN']]$, and hence, $[[us1 & us2 : Θ | uv uN]]$; 
            \item $[[us1 & us2]]$ is equivalence-only; 
            \item $[[us1 & us2]]$ is normalized;
            \item $[[us1 & us2 | uv uP]] = [[us1]]$ and hence, $[[ [us1 & us2]uP ]] = [[ [us1]uP ]] = [[iQ]]$;
            \item $[[us1 & us2 | uv uN']] = [[us2]]$ and hence, $[[ [us1 & us2]uN' ]] = [[ [us2]uN' ]] = [[iM']]$.
        \end{itemize}
        This way, $[[ [us1 & us2]uN ]] = [[ [us1 & us2]uP → [us1 & us2]uN' ]] = [[ iQ → iM' ]] = [[ iM ]]$.

@@

\begin{lemma}[Completeness of Unification] \label{lemma:unification-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] For normalized $[[uP]]$ and $[[iQ]]$ such that
        $[[Γ ; Θ ⊢ uP]]$ and $[[Γ ⊢ iQ]]$,\\ 
        assume there exists an equivalence-only unification solution
        $[[us : Θ | uv uP]]$ such that $[[ [us]uP = iQ ]]$.
        Then $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$.

        \item [$-$] For normalized $[[uN]]$ and $[[iM]]$ such that
        $[[Γ ; Θ ⊢ uN]]$ and $[[Γ ⊢ iM]]$,\\
        assume there exists an equivalence-only unification solution
        $[[us : Θ | uv uN]]$        
        such that $[[ [us]uN = iM ]]$.
        Then $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$.
   \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on the structure of $[[uP]]$ and mutually, $[[uN]]$.
    \begin{caseof}
        \item $[[uN]] = [[α̂⁻]]$\\
            Since $[[uv uN]]$ = $[[α̂⁻]]$, 
            $[[us : Θ | {α̂⁻}]]$ means that $[[ α̂⁻[Δ] ]] \in [[Θ]]$, 
            and since $[[us]]$ is equivalence only, $[[us]] = [[ (α̂⁻ :≈ iM') ]]$
            for some $[[Δ ⊢ iM']]$.
            $[[ [us]uN = iM ]]$ means $[[iM' = iM]]$.
            This way, \ruleref{\ottdruleUNUVarLabel} is applicable to infer 
            $[[Γ ; Θ ⊨ â⁻ ≈u iM ⫤ (â⁻ :≈ iM)]]$.
        \item $[[uN]] = [[α⁻]]$\\
            Then $[[us : Θ | {α⁻}]]$ means that $[[ us ]] = [[·]]$, 
            and $[[ [us]α⁻ = iM ]]$ means $[[iM = α⁻]]$.
            This way, \ruleref{\ottdruleUNVarLabel} infers 
            $[[Γ; Θ ⊨ a⁻ ≈u a⁻ ⫤ ·]]$, which is rewritten as $[[Γ; Θ ⊨ uN ≈u iM ⫤ us]]$.

@@

            Then by the induction hypothesis, $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us|uv uP]]$ and
            $[[Γ ; Θ ⊨ uN' ≈u iM' ⫤ us|uv uN']]$.
            To apply \ruleref{\ottdruleUArrowLabel} and infer the required
            $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$, we need to show that
            $[[us|uv uP & us|uv uN']] = [[us]]$. 
            By the completeness of the unification solution merge 
            (\cref{lemma:unif-sol-merge-completeness}), 
            $[[us|uv uP & us|uv uN']] = [[us|uv uP ∪ uv uN']]$,
            and $[[us|uv uP ∪ uv uN']] = [[us|uv uN]] = [[us]]$, since $[[us : Θ | uv uN]]$.

>>>>

        By the soundness of the unification solution merge (\cref{corollary:unif-sol-merge-soundness}), we have:
        \begin{itemize}
            \item $[[us1 & us2 : Θ | uv uP ∪ uv uN']]$, and hence, $[[us1 & us2 : Θ | uv uN]]$; 
            \item $[[us1 & us2]]$ is equivalence-only; 
            \item $[[us1 & us2]]$ is normalized;
            \item $[[us1 & us2 | uv uP]] = [[us1]]$ and hence, $[[ [us1 & us2]uP ]] = [[ [us1]uP ]] = [[iQ]]$;
            \item $[[us1 & us2 | uv uN']] = [[us2]]$ and hence, $[[ [us1 & us2]uN' ]] = [[ [us2]uN' ]] = [[iM']]$.
        \end{itemize}
        This way, $[[ [us1 & us2]uN ]] = [[ [us1 & us2]uP → [us1 & us2]uN' ]] = [[ iQ → iM' ]] = [[ iM ]]$.

@@

\begin{lemma}[Completeness of Unification] \label{lemma:unification-completeness}
    \hfill
    \begin{itemize}
        \item [$+$] For normalized $[[uP]]$ and $[[iQ]]$ such that
        $[[Γ ; Θ ⊢ uP]]$ and $[[Γ ⊢ iQ]]$,\\ 
        assume there exists a normalized unification solution
        $[[us : Θ | uv uP]]$ such that $[[ [us]uP = iQ ]]$.
        Then $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us]]$.

        \item [$-$] For normalized $[[uN]]$ and $[[iM]]$ such that
        $[[Γ ; Θ ⊢ uN]]$ and $[[Γ ⊢ iM]]$,\\
        assume there exists a normalized unification solution
        $[[us : Θ | uv uN]]$        
        such that $[[ [us]uN = iM ]]$.
        Then $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$.
   \end{itemize}
\end{lemma}
\begin{proof}
    We prove it by induction on the structure of $[[uP]]$ and mutually, $[[uN]]$.
    \begin{caseof}
        \item $[[uN]] = [[α̂⁻]]$\\
            Since $[[uv uN]]$ = $[[α̂⁻]]$, 
            $[[us : Θ | {α̂⁻}]]$ means that $[[ α̂⁻[Δ] ]] \in [[Θ]]$, 
            and since $[[us]]$ is equivalence only, $[[us]] = [[ (α̂⁻ :≈ iM') ]]$
            for some $[[Δ ⊢ iM']]$.
            $[[ [us]uN = iM ]]$ means $[[iM' = iM]]$.
            This way, \ruleref{\ottdruleUNUVarLabel} is applicable to infer 
            $[[Γ ; Θ ⊨ â⁻ ≈u iM ⫤ (â⁻ :≈ iM)]]$.
        \item $[[uN]] = [[α⁻]]$\\
            Then $[[us : Θ | {α⁻}]]$ means that $[[ us ]] = [[·]]$, 
            and $[[ [us]α⁻ = iM ]]$ means $[[iM = α⁻]]$.
            This way, \ruleref{\ottdruleUNVarLabel} infers 
            $[[Γ; Θ ⊨ a⁻ ≈u a⁻ ⫤ ·]]$, which is rewritten as $[[Γ; Θ ⊨ uN ≈u iM ⫤ us]]$.

@@

            Then by the induction hypothesis, $[[Γ ; Θ ⊨ uP ≈u iQ ⫤ us|uv uP]]$ and
            $[[Γ ; Θ ⊨ uN' ≈u iM' ⫤ us|uv uN']]$.
            To apply \ruleref{\ottdruleUArrowLabel} and infer the required
            $[[Γ ; Θ ⊨ uN ≈u iM ⫤ us]]$, we need to show that
            $[[us|uv uP & us|uv uN']]$ is defined and equal to $[[us]]$. 
            It holds by the completeness  of the unification solution merge 
            (\cref{lemma:unif-sol-merge-completeness}):
            \begin{itemize}
                \item $[[us|uv uP : Θ | uv uP]]$,
                \item $[[us|uv uN' : Θ | uv uN']]$,
                \item $[[us|uv uP]]$ and  $[[us|uv uN']]$ are normalized unification solution since so is $[[us]]$,
                \item $[[us : Θ | uv uP ∪ uv uN']]$ since $[[uv uP ∪ uv uN']] = [[uv uN]]$
            \end{itemize}
